Computational methods in landscape ecology

### 2024-05-21 Sections/spatial-patterns.qmd
adds a bunch of ieas

@@

intro to entropy
various concepts: thermodynamic (Boltzmann), information (Shannon), Kolmogorov, etc.
how they are related to each other
how they are related to spatial patterns and landscape ecology

@@

@@

add an info about shape metrics

>>>>

@daleSpatialPatternAnalysis2000
@wiegandUsingPatternOriented2003

@@

intro to entropy
various concepts:

- information (Shannon)/Shannon diversity
@Shannon1948
@nowosadInformationTheoryConsistent2019a

- thermodynamic (Boltzmann)

@Cushman2016
@cushmanGeneralizingBoltzmannConfigurational2021
@gaoComputationBoltzmannEntropy2019
@zhangCalculatingWassersteinMetricBased2020a
@stepinskiCurbYourEnthusiasm2022

- Kolmogorov, etc.; Juhász-Nagy’s information theory-based functions, Simpson index, Rao index, Pielou, Renyu, ...
@konradAdaptingMultiscaleApproach2023; 
@simpsonMeasurementDiversity1949

how they are related to each other
how they are related to spatial patterns and landscape ecology

@@

<!-- https://r-spatialecology.github.io/landscapemetrics/ -->
<!-- https://jakubnowosad.com/bespatial/ -->
<!-- https://cran.r-project.org/web/packages/LandComp/index.html -->

@@

add an info about shape metrics
@basaranerPerformanceShapeIndices2017

https://github.com/r-spatialecology/vectormetrics

<!--misc-->
<!-- the need for open-source software in landscape ecology -->
<!-- every proposed method should be open-source and available to the community -->

## Operations on spatial patterns

@nowosadMotifOpensourceTool2021

<!-- FOSS4G Europe pap? -->
